### 2021-11-21 draft.tex
Deploying to gh-pages from @ gottacatchenall/thesis_proposal@794d93a51a4345d2c4e99e3bb6bf91a3eb762279 🚀

Interactions between plants and pollinators create networks of
interactions, which together form the ``architecture of biodiversity''
(Bascompte \& Jordano 2007). The functioning and stability of ecosystems
emerge from these interactions, but anthropogenic change threatens to
unravel and ``rewire'' these interaction networks (CaraDonna \emph{et
al.} 2017), threatening the persistence of these systems.
Plant-pollinator networks face two possible forms of rewiring in
response to anthropogenic environmental change: spatial and temporal.
Spatially, range shifts could cause interacting species to no longer
overlap in space, and shifts in phenology could cause interacting
species to no longer occur at the same time of year. This chapter uses
several years of data on bumblebee-flower phenology and interactions
across several field sites, each consisting of several plots across an
elevational gradient, combined with spatial records of species
occurrence via GBIF to forecast this uncoupling. This addresses the EBV
to forecast element of the flow from data to mitigation in
fig.~\ref{fig:thesis} (left).

@@

This chapter uses simulation models to investigate the relationship
between species relative abundance, sampling effort, and probability of
accurately detecting an interaction between species, and further
proposes a method for optimizing the spatial sampling locations to
maximize the probability of detecting an interaction between two species
given their distributions. This addresses the optimization of monitoring
network part of the flow from data to mitigation at the top of
fig.~\ref{fig:thesis}, left. As explored in the previous chapter, there
are false-negatives in interaction data. However, there is more than one
way to observe a false-negative when sampling interactions.
fig.~\ref{fig:fnrtaxonomy} shows a taxonomy of false-negatives in
occurrence, co-occurrence, and interaction data.

@@

The first result is using a log-normal distribution of relative
abundance (Hubbell 2001) to compute a null expectation of the number of
total observations of individuals of \emph{any species} by simulating
the process of observation on food-webs generated using the niche model
(Williams \& Martinez 2000) with connectance parameterized by the
flexible-links model (MacDonald \emph{et al.} 2020). An example of this
relation for networks with varying spceies richness is shown in
fig.~\ref{fig:fnr}.

@@

We then go on to testing some assumptions of this neutral model with
empirical data. Primarily that we analytically show that our neutral
model, if anything, underestimates the probability of false-negatives if
there are positive associations between species co-occurrence, which we
show exist in two sets of spatially replicated samples of interaction
networks (Thompson \& Townsend 2000; Hadfield \emph{et al.} 2014),
fig.~\ref{fig:posassoc}---further I'm planning to add the field data
from the previous chapter into this analysis once available.

@@

Finally this chapter proposes a simulated annealing method to optimize
the efficacy of interaction detection given a set of observation
locations for a pair of species \emph{known} distributions \(D_a, D_b\).

@@

\begin{figure}
\hypertarget{fig:ch3}{%
\centering
\includegraphics{./figures/ch3.png}
\caption{fig}\label{fig:ch3}
}
\end{figure}

@@

We propose various landscape modifications which alter the cover of a
landscape, represented as a raster. We then compute a new resistance
surface based on the proposed landscape modification, and based on the
values of resistance to dispersal between each location we simulate
spatially-explicit metapopulation dynamics model (Hanski \& Ovaskainen
2000; Ovaskainen \emph{et al.} 2002) to estimate a distribution of time
until extinction for each landscape modification.

We then use simulated annealing to explore the search space of possible
modifications to estimate the modification that maximizes the time-until
extinction of simulated metapopulation dynamics under that hypothetical
modified landscape.

\begin{itemize}
\tightlist
\item
  brief overview of simulated annealing describe how you build the
\item
  proposal function optimize landscape optimization
\end{itemize}

@@

This chapter consists of a collection of modules in the Julia language
for different aspects of community ecology, including most of the code
used for the preceding chapters. Indeed
\texttt{MetacommunityDynamics.jl} (MCD.jl) is the epicenter of this set
of tools, but due to the nature of the Julia language, MCD.jl is
interoperable with serveral existing packages within the
\texttt{EcoJulia} organization, including several to which I have
contributed. A diagram showing the relation between these packages is
shown in fig.~\ref{fig:software}.

@@

>>>>

Interactions between plants and pollinators create networks of
interactions, which together form the ``architecture of biodiversity''
(Bascompte \& Jordano 2007). The functioning and stability of ecosystems
emerge from these interactions, but anthropogenic change threatens to
unravel and ``rewire'' these interaction networks (CaraDonna \emph{et
al.} 2017), threatening the persistence of these systems.
Plant-pollinator networks face two possible forms of rewiring in
response to anthropogenic environmental change: spatial and temporal.
Spatially, range shifts could cause interacting species to no longer
overlap in space, and shifts in phenology could cause interacting
species to no longer occur at the same time of year. This chapter uses
several years of data on bumblebee-flower phenology and interactions
across several field sites, each consisting of several plots across an
elevational gradient, combined with spatial records of species
occurrence via GBIF to forecast the uncoupling of the plant-pollinator
metaweb of Colorado.

@@

This chapter uses simulation models to investigate the relationship
between species relative abundance, sampling effort, and probability of
observing an interaction between species, and further proposes a method
for optimizing the spatial sampling locations to maximize the
probability of detecting an interaction between two species given their
distributions. This addresses the optimization of monitoring network
part of the flow from data to mitigation at the top of
fig.~\ref{fig:thesis}, left. As explored in the previous chapter, there
are false-negatives in interaction data. However, there is more than one
way to observe a false-negative when sampling interactions.
fig.~\ref{fig:fnrtaxonomy} shows a taxonomy of false-negatives in
occurrence, co-occurrence, and interaction data.

@@

The first result is to compute a null expectation of the probability of
an interaction false-negative as a function number of total observations
of individuals of \emph{any species}. This is done by using a log-normal
distribution of relative abundance (Hubbell 2001) and simulating the
process of observation on food-webs generated using the niche model
(Williams \& Martinez 2000) with connectance parameterized by the
flexible-links model (MacDonald \emph{et al.} 2020). An example of this
relation for networks with varying spceies richness is shown in
fig.~\ref{fig:fnr}.

@@

We then go on to testing some assumptions of this neutral model with
empirical data. Primarily that we analytically show that our neutral
model, if anything, underestimates the probability of false-negatives if
there are positive associations between species co-occurrence, and we
show these positive associations exist in two sets of spatially
replicated samples of interaction networks (Thompson \& Townsend 2000;
Hadfield \emph{et al.} 2014), fig.~\ref{fig:posassoc}---further I'm
planning to add the field data from the previous chapter into this
analysis once available.

@@

Finally this chapter proposes a simulated annealing method to optimize
the a set of \(n\) points in space to maximize the probability of
detecting an interaction between two species \(a\) and \(b\) with
\emph{known} distributions \(D_a\), \(D_b\).

@@

\begin{figure}
\hypertarget{fig:ch3}{%
\centering
\includegraphics{./figures/ch3.png}
\caption{foo}\label{fig:ch3}
}
\end{figure}

@@

We propose various landscape modifications which alter the cover of a
landscape, represented as a raster. We then compute a new resistance
surface based on the proposed landscape modification using Circuitscape
(McRae \emph{et al.} 2008), and based on the values of resistance to
dispersal between pair of locations we simulate spatially-explicit
metapopulation dynamics model (Hanski \& Ovaskainen 2000; Ovaskainen
\emph{et al.} 2002) to estimate a distribution of time until extinction
for each landscape modification. The largest challenge in implementing
this algorithm is the space of potential modifications grows as
\(O(nm)\) for an \(n\) by \(m\) raster. For most actual landscapes to
which we wish to apply this method, the set of possible modifications
becomes uncomputably large, so we use simulated annealing to explore the
search space of possible modifications to estimate the modification that
maximizes the time-until extinction of simulated metapopulation dynamics
under that hypothetical modified landscape.

The biggest challenge in implementing simulated annealing in this
context is defining a proposal function for landscape modifications.
This is done by computing the minimum-spanning-tree (MST) of the spatial
nodes, and then proposing corridors that connect nodes that are already
connected in the MST.

@@

The final chapter consists of a collection of modules in the Julia
language for different aspects of community ecology, including most of
the code used for the preceding chapters. Indeed
\texttt{MetacommunityDynamics.jl} (MCD.jl) is the epicenter of this set
of tools, but due to the nature of the Julia language, MCD.jl is
interoperable with serveral existing packages within the
\texttt{EcoJulia} organization, including several to which I have
contributed. A diagram showing the relation between these packages is
shown in fig.~\ref{fig:software}.

@@

\leavevmode\hypertarget{ref-McRae2008UsiCir}{}%
McRae, B.H., Dickson, B.G., Keitt, T.H. \& Shah, V.B. (2008). Using
Circuit Theory to Model Connectivity in Ecology, Evolution, and
Conservation. \emph{Ecology}, 89, 2712--2724.
